An excess of positive results: Comparing the standard Psychology literature with Registered Reports

### 2020-01-13 preprint/manuscript.Rmd
minor changes to explanations of Figure 3

# Discussion
We examined the proportion of Psychology articles that find support for their first tested hypothesis and found a large difference ($`r printnum(prop.support.SR*100)` \%$ vs $`r printnum(prop.support.RR*100)` \%$) between a random sample of standard reports and the full population of Registered Reports (at the time of data analysis). 
More than half of the analysed Registered Reports tested hypotheses based on direct replications, but the difference between standard reports and Registered Reports was still large when all direct replications were excluded from the analysis ($`r printnum(prop.orig.support.SR*100)` \%$ vs $`r printnum(prop.orig.support.RR*100)` \%$). 
The introduction of Registered Reports has clearly led to a much larger proportion of null results appearing in the published literature compared to standard reports.

@@

To explain the $`r printnum((prop.support.SR - prop.support.RR)*100)` \%$ gap between standard reports and Registered Reports, we must assume some combination of differences in bias, statistical power, or the proportion of true hypotheses researchers choose to examine. 
Figure\ \@ref(fig:powerbaserate) visualises the combinations of statistical power and proportion of true hypotheses that would produce the observed positive result rates if the literature were completely unbiased. 
For example, if we assume no publication bias and no QRPs, more than $90\%$ of the hypotheses authors of standard reports study must be true and study designs must have more than $90\%$ power for the true effect size. 
This is highly unlikely, meaning that the standard literature is unlikely to reflect reality. 
As we noted above, there is good reason to assume that methodological rigour and statistical power in Registered Reports are as high, or higher, than in standard reports, leaving the rate of true hypotheses and bias as remaining explanations.

(ref:powerbaserate) Combinations of the proportion of true hypotheses and statistical power that would produce the observed positive result rates given $\alpha = 5 \%$ and no bias. Shaded areas indicate $95\%$ confidence intervals. SRs = standard reports, RRs = Registered Reports. The curve for all SRs (i.e, including replications; $`r printnum(prop.support.SR*100)` \%$ positive results, $N = `r n.SR`$) is not shown because it is almost identical to the one for original SRs. Plotted values were calculated using the equation $PRR = \alpha*(1-t) + (1-\beta)*t$; with $PRR =$ positive result rate, $\alpha =$ probability of obtaining a positive result when testing a false hypothesis (here fixed at .05), $1-\beta =$ probability of obtaining a positive result when testing a true hypothesis (power), and $t =$ proportion of true hypotheses; and solving for $t$ and $1-\beta$, respectively.

@@

## Limitations
We compared hypotheses tested in Registered Reports with hypotheses tested in standard reports. 
Because neither hypotheses, authors, nor editors were randomly assigned to each publication format we cannot draw firm conclusions about the causes that led to a difference in the proportion of supported hypotheses. 
Although it seems plausible that selective reporting and QRPs are reduced in Registered Reports, we do not know by how much, nor if this reduction would be of comparable size in a randomised experiment. 
As mentioned above, it is a-priori plausible that the Registered Reports format is used selectively for particularly risky hypotheses. 
This means that the proportion of true hypotheses in Registered Reports does not necessarily generalise to the entire population of hypotheses that are tested in Psychology. 
It is also important to note that our results do not warrant the conclusion that Registered Reports are effective at reducing all forms of bias. 
Authors self-select to submit Registered Reports, and the format may be particularly popular among those who try to minimise the risk of inflated error rates regardless of the report format they use. 
This would lead to less bias in the Registered Reports literature even if the format's safeguards against certain QRPs were actually ineffective.

>>>>

# Discussion
We examined the proportion of Psychology articles that find support for their first tested hypothesis and discovered a large difference ($`r printnum(prop.support.SR*100)` \%$ vs $`r printnum(prop.support.RR*100)` \%$) between a random sample of standard reports and the full population of Registered Reports (at the time of data analysis). 
More than half of the analysed hypothesis tests in Registered Reports were direct replications of previous work, but the difference between standard reports and Registered Reports was still large when direct replications were excluded from the analysis ($`r printnum(prop.orig.support.SR*100)` \%$ vs $`r printnum(prop.orig.support.RR*100)` \%$). 
The introduction of Registered Reports has clearly led to a much larger proportion of null results appearing in the published literature compared to standard reports.

@@

To explain the $`r printnum((prop.support.SR - prop.support.RR)*100)` \%$ gap between standard reports and Registered Reports, we must assume some combination of differences in bias, statistical power, or the proportion of true hypotheses researchers choose to examine. 
Figure\ \@ref(fig:powerbaserate) visualises the combinations of statistical power and proportion of true hypotheses that would produce the observed positive result rates if the literature were completely unbiased. 
For example, if we assume no publication bias and no QRPs, more than $90\%$ of the hypotheses authors of standard reports study must be true and study designs must have more than $90\%$ power for the true effect size. 
This is highly unlikely, meaning that the standard literature is unlikely to reflect reality. 
As we noted above, there is good reason to assume that methodological rigour and statistical power in Registered Reports are as high as in standard reports or higher, leaving the rate of true hypotheses and bias as remaining explanations.

(ref:powerbaserate) Combinations of the proportion of true hypotheses and statistical power that would produce the observed positive result rates given $\alpha = 5 \%$ and no bias. Shaded areas indicate $95\%$ confidence intervals. SRs = standard reports, RRs = Registered Reports. The curve for all SRs (i.e, including replications; $`r printnum(prop.support.SR*100)` \%$ positive results, $N = `r n.SR`$) is not shown because it is almost identical to the one for original SRs. Plotted values were calculated using the equation $PRR = \alpha*(1-t) + (1-\beta)*t$; with $PRR =$ positive result rate, $\alpha =$ probability of obtaining a positive result when testing a false hypothesis (here fixed at .05), $1-\beta =$ probability of obtaining a positive result when testing a true hypothesis (power), and $t =$ proportion of true hypotheses; and solving for $t$ and $1-\beta$, respectively (with the simplifying assumption that all studies in one group have the same power).

@@

## Limitations
We compared hypotheses tested in Registered Reports with hypotheses tested in standard reports. 
Because neither hypotheses, authors, nor editors were randomly assigned to each publication format, we cannot draw firm conclusions about the causes that led to a difference in the proportion of supported hypotheses. 
Although it seems plausible that selective reporting and QRPs are reduced in Registered Reports, we do not know by how much, nor if this reduction would be of comparable size in a randomised experiment. 
As mentioned above, it is a-priori plausible that the Registered Reports format is used selectively for particularly risky hypotheses. 
This means that the proportion of true hypotheses in Registered Reports does not necessarily generalise to the entire population of hypotheses that are tested in Psychology. 
It is also important to note that our results do not warrant the conclusion that Registered Reports are effective at reducing all forms of bias. 
Authors self-select to submit Registered Reports, and the format may be particularly popular among those who try to minimise the risk of inflated error rates regardless of the report format they use. 
This would lead to less bias in the Registered Reports literature even if the format's safeguards against certain QRPs were actually ineffective.

### 2020-01-13 preprint/manuscript.tex
minor changes to explanations of Figure 3

We examined the proportion of Psychology articles that find support for their first tested hypothesis and found a large difference (\(96.05 \%\) vs \(43.66 \%\)) between a random sample of standard reports and the full population of Registered Reports (at the time of data analysis).
More than half of the analysed Registered Reports tested hypotheses based on direct replications, but the difference between standard reports and Registered Reports was still large when all direct replications were excluded from the analysis (\(95.95 \%\) vs \(50.00 \%\)).
The introduction of Registered Reports has clearly led to a much larger proportion of null results appearing in the published literature compared to standard reports.

@@

To explain the \(52.39 \%\) gap between standard reports and Registered Reports, we must assume some combination of differences in bias, statistical power, or the proportion of true hypotheses researchers choose to examine.
Figure~\ref{fig:powerbaserate} visualises the combinations of statistical power and proportion of true hypotheses that would produce the observed positive result rates if the literature were completely unbiased.
For example, if we assume no publication bias and no QRPs, more than \(90\%\) of the hypotheses authors of standard reports study must be true and study designs must have more than \(90\%\) power for the true effect size.
This is highly unlikely, meaning that the standard literature is unlikely to reflect reality.
As we noted above, there is good reason to assume that methodological rigour and statistical power in Registered Reports are as high, or higher, than in standard reports, leaving the rate of true hypotheses and bias as remaining explanations.

@@

\caption{Combinations of the proportion of true hypotheses and statistical power that would produce the observed positive result rates given \(\alpha = 5 \%\) and no bias. Shaded areas indicate \(95\%\) confidence intervals. SRs = standard reports, RRs = Registered Reports. The curve for all SRs (i.e, including replications; \(96.05 \%\) positive results, \(N = 152\)) is not shown because it is almost identical to the one for original SRs. Plotted values were calculated using the equation \(PRR = \alpha*(1-t) + (1-\beta)*t\); with \(PRR =\) positive result rate, \(\alpha =\) probability of obtaining a positive result when testing a false hypothesis (here fixed at .05), \(1-\beta =\) probability of obtaining a positive result when testing a true hypothesis (power), and \(t =\) proportion of true hypotheses; and solving for \(t\) and \(1-\beta\), respectively.}\label{fig:powerbaserate}
\end{figure*}

@@

We compared hypotheses tested in Registered Reports with hypotheses tested in standard reports.
Because neither hypotheses, authors, nor editors were randomly assigned to each publication format we cannot draw firm conclusions about the causes that led to a difference in the proportion of supported hypotheses.
Although it seems plausible that selective reporting and QRPs are reduced in Registered Reports, we do not know by how much, nor if this reduction would be of comparable size in a randomised experiment.
As mentioned above, it is a-priori plausible that the Registered Reports format is used selectively for particularly risky hypotheses.
This means that the proportion of true hypotheses in Registered Reports does not necessarily generalise to the entire population of hypotheses that are tested in Psychology.
It is also important to note that our results do not warrant the conclusion that Registered Reports are effective at reducing all forms of bias.
Authors self-select to submit Registered Reports, and the format may be particularly popular among those who try to minimise the risk of inflated error rates regardless of the report format they use.
This would lead to less bias in the Registered Reports literature even if the format's safeguards against certain QRPs were actually ineffective.

>>>>

We examined the proportion of Psychology articles that find support for their first tested hypothesis and discovered a large difference (\(96.05 \%\) vs \(43.66 \%\)) between a random sample of standard reports and the full population of Registered Reports (at the time of data analysis).
More than half of the analysed hypothesis tests in Registered Reports were direct replications of previous work, but the difference between standard reports and Registered Reports was still large when direct replications were excluded from the analysis (\(95.95 \%\) vs \(50.00 \%\)).
The introduction of Registered Reports has clearly led to a much larger proportion of null results appearing in the published literature compared to standard reports.

@@

To explain the \(52.39 \%\) gap between standard reports and Registered Reports, we must assume some combination of differences in bias, statistical power, or the proportion of true hypotheses researchers choose to examine.
Figure~\ref{fig:powerbaserate} visualises the combinations of statistical power and proportion of true hypotheses that would produce the observed positive result rates if the literature were completely unbiased.
For example, if we assume no publication bias and no QRPs, more than \(90\%\) of the hypotheses authors of standard reports study must be true and study designs must have more than \(90\%\) power for the true effect size.
This is highly unlikely, meaning that the standard literature is unlikely to reflect reality.
As we noted above, there is good reason to assume that methodological rigour and statistical power in Registered Reports are as high as in standard reports or higher, leaving the rate of true hypotheses and bias as remaining explanations.

@@

\caption{Combinations of the proportion of true hypotheses and statistical power that would produce the observed positive result rates given \(\alpha = 5 \%\) and no bias. Shaded areas indicate \(95\%\) confidence intervals. SRs = standard reports, RRs = Registered Reports. The curve for all SRs (i.e, including replications; \(96.05 \%\) positive results, \(N = 152\)) is not shown because it is almost identical to the one for original SRs. Plotted values were calculated using the equation \(PRR = \alpha*(1-t) + (1-\beta)*t\); with \(PRR =\) positive result rate, \(\alpha =\) probability of obtaining a positive result when testing a false hypothesis (here fixed at .05), \(1-\beta =\) probability of obtaining a positive result when testing a true hypothesis (power), and \(t =\) proportion of true hypotheses; and solving for \(t\) and \(1-\beta\), respectively (with the simplifying assumption that all studies in one group have the same power).}\label{fig:powerbaserate}
\end{figure*}

@@

We compared hypotheses tested in Registered Reports with hypotheses tested in standard reports.
Because neither hypotheses, authors, nor editors were randomly assigned to each publication format, we cannot draw firm conclusions about the causes that led to a difference in the proportion of supported hypotheses.
Although it seems plausible that selective reporting and QRPs are reduced in Registered Reports, we do not know by how much, nor if this reduction would be of comparable size in a randomised experiment.
As mentioned above, it is a-priori plausible that the Registered Reports format is used selectively for particularly risky hypotheses.
This means that the proportion of true hypotheses in Registered Reports does not necessarily generalise to the entire population of hypotheses that are tested in Psychology.
It is also important to note that our results do not warrant the conclusion that Registered Reports are effective at reducing all forms of bias.
Authors self-select to submit Registered Reports, and the format may be particularly popular among those who try to minimise the risk of inflated error rates regardless of the report format they use.
This would lead to less bias in the Registered Reports literature even if the format's safeguards against certain QRPs were actually ineffective.
